---
title: "Covid91 vaccine study Final2024"
author: "Ashia Graves!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "December 2024`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males <- guy %>%
  group_by(infected,treatment) %>%
  summarize(count=n()) %>% 
  mutate(prcnt=count/sum(count))

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of males people and the relationship of covid-19 by either taking the vaccine or the placebo. 


The plot shows covid19 being very low with those that took the vaccine in comparison to those with the placebo. The Placebo compared to the vaccine is about 3:1 in size.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

Only 26.2% of males who received the vaccine contracted COVID-19, whereas 73.8% of those who did not receive the vaccine were infected. This suggests that the vaccine shows great promise.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```


The key takeaway from these results is the outcome of Fisher's Exact test. The p-value of 2.2e-16 (0.00000000000000022) is exceptionally small, indicating that these results are highly unlikely to be due to chance.

Additionally, our odds ratio of approximately 2.8 demonstrates that the vaccine is effective in preventing COVID-19 when compared to a placebo.


## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions