---
title: "Covid91 vaccine study Final2024"
author: "Ashia Graves!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---
 
```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males <- guy %>%
  group_by(infected,treatment) %>%
  summarize(count=n()) %>% 
  mutate(prcnt=count/sum(count))

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of males people and the relationship of covid-19 by either taking the vaccine or the placebo. 


The plot shows covid19 being very low with those that took the vaccine in comparison to those with the placebo. The Placebo compared to the vaccine is about 3:1 in size.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

Only 26.2% of males who received the vaccine contracted COVID-19, whereas 73.8% of those who did not receive the vaccine were infected. This suggests that the vaccine shows great promise.


### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```


The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 2.2e-16  0.00000000000000022. Seeing this number means that these results are not likely to occur by chance.

Our odds ratio of about 2.8 proves that this vaccine is effective in stopping covid19 compared to a placebo.
```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```

### Effectiveness Confidence Interval `r low`  to `r hi`


The odds ratio of the males with vaccines vs those with placebos has a ratio greater than 70 meaning we have a effective vaccine for males. 


### Barplot

```{r}
Females <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot illustrates the relationship between COVID-19 infection rates in females and whether they received the vaccine or the placebo.

The data shows a 2:1 ratio of infection rates between the vaccine and placebo groups. Approximately 36% of females who received the vaccine contracted COVID-19, compared to around 65% of females who received the placebo. 

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```
 
### Numerical results
These numerical results show specific percentages of drugusers that got covid-19 after taking the placebo and taking the vaccine.
Drugusers had a percentage of 91.7% having the vaccine who got infected. Compared to those with a placebo who had 8.3% got infected. This looks like a horrible vaccine for drugusers to take.

### Inferential 




```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```


The p-value is 0.00000007. This in an incredibly small number and with that being said, the conclusion is that these results show that something is happening.

The Fishers's Exact Test illustrates that the odds ratio is 4.3
Proving that this vaccine is very effective.

```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 for those who identified as gay.  
The odds ratio is 81 which is the highest we have seen in all data points so far. This is a extremely good vaccine for those that are gay.

### Barplot


```{r}
Druggies <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of people who answered  Druguser to drug use and the relationship to the contraction of covid-19 while taking either the vaccine or the placebo.

Those who are drug users who took the vaccine had about 80% infection compared to those who took a placebo had about 8% infected. This is a ration of 1:10. 

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```
 

These numerical results show specific percentages of drugusers that got covid-19 after taking the placebo and taking the vaccine.
Drugusers had a percentage of 91.7% having the vaccine who got infected. Compared to those with a placebo who had 8.3% got infected. This looks like a horrible vaccine for drugusers to take.
### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```


The p-value is 0.000000000003. This in a small number and with that being said, Meaaing something is happening in this graph.

The Fishers's Exact Test illustrates that the odds ratio is 0.087
This shows that the vaccine is horrible and should never be used for drug users especially with how small the odds ratio is.

```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```

```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`




Above, is the effectiveness for the covid-19 for guys. If the effectiveness is above 50% that means that the vaccine is effective. The confidence interval odds ratio shows to be an 8, this is less than 50 and this is very very small meaning this vaccine should never be conssidered and it is worse that taking nothing as it gives covid19 and taking nothing is better than this vaccine.



# Overall Results and Conclusions
In conclusion, the effectiveness of the COVID vaccine varies across different subgroups. Starting with the male group, we reject the null hypothesis. Based on the bar chart, numerical data, and the results of the chi-squared and Fisher’s Exact tests, we can conclude that men who received the COVID vaccine and contracted the virus experienced fewer symptoms compared to those who received the placebo. For men who were unvaccinated and did not contract COVID, there was little to no difference between the groups. The infection rate in the placebo group was 2.55%, while the vaccinated group had an infection rate of 1.15%, suggesting the vaccine’s effectiveness in men. Therefore, we reject the null hypothesis for the male subgroup.

For the female group, we also reject the null hypothesis. All tests indicate that women who received the COVID vaccine and contracted the virus experienced fewer symptoms than those who received the placebo. The unvaccinated, non-infected women showed little to no difference compared to each other. In the placebo group, the infection rate was 2.04%, while the vaccinated group had an infection rate of 1.37%, further supporting the vaccine’s effectiveness in women. Thus, we reject the null hypothesis for the female subgroup as well.

When comparing the male and female subgroups, women had slightly lower infection rates with both the COVID vaccine and placebo, by 0.22% and 0.51%, respectively. While this difference is minimal, the data suggests that women experience fewer symptoms with the vaccine.

In the LGBTQ subgroup, we also reject the null hypothesis, but with different results. The data shows that the vaccine had a negative effect in this group. LGBTQ individuals who received the COVID vaccine and contracted the virus experienced more severe symptoms than those who received the placebo. The non-vaccinated, non-infected LGBTQ group showed little to no difference between the two groups. The placebo group had an infection rate of 0.66%, while the vaccinated group had an infection rate of 3.54%, a significant difference. Therefore, we reject the null hypothesis, as the vaccine had an impact on this subgroup, though negative.

For the drug user subgroup, the results mirror those of the LGBTQ subgroup. We reject the null hypothesis, as drug users who received the COVID vaccine and contracted the virus experienced significantly more symptoms compared to those who received the placebo. The vaccinated drug user group had an infection rate of 3.06%, while the placebo group had an infection rate of 0.4%, showing a substantial difference in the severity of symptoms. Thus, we reject the null hypothesis in this case as well.

When comparing all the subgroups based on the effects of the COVID vaccine, the female group showed the best results, experiencing fewer symptoms with the vaccine. The LGBTQ subgroup, on the other hand, had the worst outcomes, experiencing the most severe symptoms despite receiving the vaccine. The male group had results similar to the female group, though slightly less favorable, and drug users experienced only marginally fewer symptoms than the LGBTQ group (by 0.48%). Based on these findings, I would recommend the COVID vaccine for males and females who are not drug users or LGBTQ, as it had a more positive effect in these groups.
