---
title: "Covid91 vaccine study Final2024"
author: "Ashia Graves!"
date: "December 2024`"
output:
  pdf_document:
    toc: true
    toc_depth: '5'
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
  word_document:
    toc: true
    toc_depth: '5'
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males <- guy %>%
  group_by(infected,treatment) %>%
  summarize(count=n()) %>% 
  mutate(prcnt=count/sum(count))

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of males people and the relationship of covid-19 by either taking the vaccine or the placebo. 


The plot shows covid19 being very low with those that took the vaccine in comparison to those with the placebo. The Placebo compared to the vaccine is about 3:1 in size.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

Only 26.2% of males who received the vaccine contracted COVID-19, whereas 73.8% of those who did not receive the vaccine were infected. This suggests that the vaccine shows great promise.


### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 2.2e-16  0.00000000000000022. Seeing this number means that these results are not likely to occur by chance.

Our odds ratio of about 2.8 proves that this vaccine is effective in stopping covid19 compared to a placebo.



### Effectiveness Confidence Interval `r low`  to `r hi`
```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)100
print(effectivenessguy)
```
The odds ratio of the males with vaccines vs those with placebos has a ratio greater than 70 meaning we have a effective vaccine for males. 


### Barplot

```{r}
Females <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot illustrates the relationship between COVID-19 infection rates in females and whether they received the vaccine or the placebo.

The data shows a 2:1 ratio of infection rates between the vaccine and placebo groups. Approximately 36% of females who received the vaccine contracted COVID-19, compared to around 65% of females who received the placebo. 

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```
 
### Numerical results
These numerical results show specific percentages of drugusers that got covid-19 after taking the placebo and taking the vaccine.
Drugusers had a percentage of 91.7% having the vaccine who got infected. Compared to those with a placebo who had 8.3% got infected. This looks like a horrible vaccine for drugusers to take.

### Inferential 




```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```


The p-value is 0.00000007. This in an incredibly small number and with that being said, the conclusion is that these results show that something is happening.

The Fishers's Exact Test illustrates that the odds ratio is 4.3
Proving that this vaccine is very effective.

```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 for those who identified as gay.  
The odds ratio is 81 which is the highest we have seen in all data points so far. This is a extremely good vaccine for those that are gay.

### Barplot


```{r}
Druggies <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of people who answered  Druguser to drug use and the relationship to the contraction of covid-19 while taking either the vaccine or the placebo.

Those who are drug users who took the vaccine had about 80% infection compared to those who took a placebo had about 8% infected. This is a ration of 1:10. 









## LGBTQ



## Druggies


# Overall Results and Conclusions